---
title: "Assignment 1 - Machine Learning Project"
author: "SiCheng Zhao"
date: '2022-05-31'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import required library
```{r}
library(iClusterPlus) 
library(GenomicRanges) 
library(gplots) 
library(lattice) 
```

## Import data set
```{r}
#setwd("/Users/scz/Documents/AI Medical and Medician /Trimiter3/Precision Medicine")
data <- read.csv(file = 'brca_data_w_subtypes.csv')
# In order to run through program, I have to use a slice of the whole data set.
GE = as.matrix(data[, 0:4])
CNA = as.matrix(data[, 605:609])
SM = as.matrix(data[, 1465:1469])
PE = as.matrix(data[, 1714:1718])
status = as.matrix(data[, (length(data)-4) : length(data)])
status[status == 'Performed but Not Available'] <- ""

new_data = as.data.frame(cbind(GE, CNA, SM, PE, status))
```

## Use the result tuned by program 
```{r}
setwd("/home/taserface/Documents/Precision Medicine ML project")
output2=alist() 
files=grep("cv2.fit",dir()) 
for(i in 1:length(files)){ 
  load(dir()[files[i]])
  output2[[i]]=cv2.fit 
} 
nLambda = nrow(output2[[1]]$lambda) 
nK = length(output2) 
BIC = getBIC(output2) 
devR = getDevR(output2)

minBICid = apply(BIC,2,which.min) 
devRatMinBIC = rep(NA,nK) 
for(i in 1:nK){ 
  devRatMinBIC[i] = devR[minBICid[i],i] 
}

plot(1:(nK+1),c(0,devRatMinBIC),type="b",xlab="Number of clusters (K+1)", 
     ylab="%Explained Variation")
```
The optimal number of latent variables is where the curve of %Explained variation levels off. By
examining the plot shown above, three is considered to be the optimal number.

```{r}
clusters  = output2[[2]]$fit[[306]]$clusters
new_data['cluster'] = clusters
```

Difference in PR status
```{r}
# PR status
counts_pr = table(new_data$cluster, new_data$PR.Status)
barplot(counts_pr, main="PR.Statis",
        col=c("darkblue","red", "yellow"),
        legend = rownames(counts_pr), beside=TRUE)

```
Difference in ER status
```{r}
# ER status
counts_er = table(new_data$cluster, new_data$ER.Status)
barplot(counts_er, main="ER.Statis",
        col=c("darkblue","red", "yellow"),
        legend = rownames(counts_er), beside=TRUE)
```
Difference in HER2 status
```{r}
# HER2 status
counts_her2 = table(new_data$cluster, new_data$HER2.Final.Status)
barplot(counts_her2, main="HER2.Statis",
        col=c("darkblue","red", "yellow"),
        legend = rownames(counts_her2), beside=TRUE)
```
Difference in histological type
```{r}
counts_his = table(new_data$cluster, new_data$histological.type)
barplot(counts_his, main="HER2.Statis",
        col=c("darkblue","red", "yellow"),
        legend = rownames(counts_his), beside=TRUE)

```

Conclusion: 